A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals

### 2019-08-29 paper/paper.intro.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

On-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals. Hospital departments provide services where patient needs, and thus the system's demands, often exceed the available supply. For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients. Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref]. It is common practice for on-call schedules to be created manually. Yet manually-created schedules are prone to errors and potential for biases [ref]. First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide an optimal schedule. Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as preventing many consecutive work blocks from being assigned or ensuring clinicians are allocated a specific amount of work blocks throughout the year. Third, manual scheduling is often time-consuming for the person developing the schedule. For these reasons, it is important to develop automated methods that can generate optimal schedules that satisfy the given constraints of the hospital department. \\

@@

Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation. While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals. % explain why hard to adapt?... what makes their generalizablility/adaptability limited?
Moreover, the large number of variables and constraints also leads to computational complexity issues \cite{goos_complexity_1996}, especially when using exact methods for finding the solution. In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St. Michael's Hospital in Toronto, Canada. Our goal is to (1) present a simple integer linear programming formulations for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem. \\
% present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments. \\ %make #3 an objective

>>>>

Hospital departments provide services where patient needs, and thus the system's demands, often exceed the available supply. In particular, on-call schedules for a fixed number of health-care providers are central to the efficient running of hospitals. Carefully allocated schedules are meant to simultaneously ensure sufficient resources are provided to patients while not overworking clinicians to prevent costly mistakes [ref]. It is common practice for on-call schedules in hospitals to be created manually, yet manually-created schedules are prone to errors and potential for various biases [ref]. First, when there is a large number of clinicians in a single department, or the constraints that need to be satisfied by the department are very complex, a manual method may not provide an optimal schedule. Second, such methods are likely to overlook certain constraints that must be maintained to have an operational department, such as preventing many consecutive work blocks from being assigned or ensuring clinicians are allocated a specific amount of work blocks throughout the year. Third, manual scheduling is often time-consuming for the person developing the schedule. For these reasons, it is important to develop automated methods that can generate optimal schedules that satisfy the given constraints of the hospital department. \\
%For example, it is important that a hospital department allocates its resources, such as the availability of a finite number of clinicians, optimally, to ensure the best possible service for its patients.

@@

Many of the approaches to nurse scheduling were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation. While these department-specific approaches allow end-users to find precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians in that department, they are difficult to readily adapt to other departments in the same hospital or other hospitals. % explain why hard to adapt?... what makes their generalizablility/adaptability limited?
Moreover, the large number of variables and constraints also leads to computational complexity issues \cite{goos_complexity_1996}, especially when using exact methods for finding the solution. In this paper, we tackle a version of the nurse scheduling problem arising from a case study of one clinical division, providing two different services simultaneously (general infectious disease (ID) consults; and HIV consults service) at St. Michael's Hospital in Toronto, Canada. Our goal is to (1) present a simple integer linear programming formulations for the scheduling problem as developed for the hospital, and describe the adaptability of the formulation to solving similar problems in other departments; (2) compare the performance of the ILP scheduler to the results of the manual approach; and (3) analyze the robustness of the algorithm in difficult instances of the problem. \\
% present a simple formulation for the problem developed and tested at the hospital after switching from a manual approach to scheduling; and (2) analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach; and (3) describe the adaptability of the formulation as a basic framework for solving similar problems in other departments. \\ %make #3 an objective

### 2019-08-29 paper/paper.methods.tex
Work on intro + references
Add majority of references for automated methods/scheduling papers in intro
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

In this section, we present an application of 0-1 Linear Programming to solve the clinician scheduling problem presented in Section \ref{sec:problem}. First, we describe the sets, indices and variables present in the formulation of the program. We then convert the constraints given in Table \ref{tbl:constraint-summary} into mathematical terms, and outline the objective function of the linear program.

@@

\subsection{Variables} \label{sec:meth-variables}
Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as $X_{c, b, s}$. A value of 1 indicates that the given clinician $c$ is assigned to provide service $s$ during block $b$. Similarly, we define weekend assignments as $Y_{c, w}$, although these are not indexed by service, as clinicians are expected to provide all services during the weekends. To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends. To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, s}$, with additional constraints on its range. Further details regarding this variable are described in Section \ref{sec:meth-objectives}. We then introduce a set of constants $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year. Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.

\begin{table}[h]
	\centering
	\begin{tabular}{ c l }
		\hline
		\textbf{Name}              & \textbf{Description}                                               \\ \hline
		$X_{c, b, s} \in \{0, 1\}$ & assignment of clinician $c$ to service $s$ on block $b$            \\
		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                         \\
		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency             \\
		$m_{c, s}$                 & minimum number of blocks clinician $c$ should cover on service $s$ \\
		$M_{c, s}$                 & maximum number of blocks clinician $c$ should cover on service $s$
	\end{tabular}
	\caption{Description of variables and constants in the problem}
	\label{tbl:variables-constants}
\end{table}

\subsection{Constraints} \label{sec:meth-constraints}
We now present the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.

\begin{align}
	&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \tag{Block Coverage} \label{eqn:constr-block-cov} \\
	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{Weekend Coverage} \label{eqn:constr-weekend-cov} \\
	&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\ c\in\mathcal{C}, s\in\mathcal{S} \tag{Min/Max} \label{eqn:constr-min-max} \\
	&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, s\in\mathcal{S} \tag{No Consecutive Blocks} \label{eqn:constr-no-consec-blocks} \\
	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{No Consecutive Weekends} \label{eqn:constr-no-consec-weekends} \\
	&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}} &&\forall c\in\mathcal{C} \tag{Equal Weekends} \label{eqn:constr-equal-weekends} \\
	&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq \ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{Equal Holidays} \label{eqn:constr-equal-holidays}
\end{align}

\subsection{Objectives} \label{sec:meth-objectives}
As described in Section \ref{sec:problem}, the soft constraints of the clinician scheduling problem include: (1) satisfying clinician block off requests, (2) satisfying clinician weekend off requests, and (3) assigning weekends closer to blocks. We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{sec:meth-variables}. Objectives (1) and (2) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to. 
\begin{align}
	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{Block Requests} \label{eqn:obj-block-requests}\\
	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{Weekend Requests} \label{eqn:obj-weekend-requests}
\end{align}
where $\ind(\cdot)$ is the indicator function that has value 1 when the condition is met and 0 otherwise. In the above two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments. The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$. For instance, clinicians might want to be assigned during a weekend that falls within an assigned block. So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5. In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned. If at least one of the two variables is not assigned, the product will be 0. This leads to the maximization objective  %avoid editorial adjectives
\begin{align}
	&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c, w=\varphi(b)} \tag{Block-Weekend Adjacency} \label{eqn:obj-block-weekend-adj}
\end{align}
where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency. For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\
% I really liked the example explanation!

However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework. An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints [ref \ref{???}]. In our case, introducing a variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that 
\begin{align}
	&Z_{c, b, s} \leq X_{c, b, s} \\
	&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S}
\end{align}
allows us to rewrite $Q_3$ as a linear function of $Z$,
\begin{align}
	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} Z_{c, b, s}
\end{align}
Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective. \\

Thus, our clinician scheduling problem is a multiple objective optimization problem. The most common approach to solving multiple objective optimization problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}]. This is the approach we decided to use in our problem, to ensure all three objectives are considered when finding a solution. Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{S}$, the normalized objectives can be written as follows,
\begin{align}
	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
\end{align}
where we divide each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
\begin{equation}
	\alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
\end{equation}
with $0 \leq \alpha, \beta \leq 1$. \\

[...]

>>>>

In this section, we present an application of integer linear programming to solve the clinician scheduling problem presented in Section \ref{sec:problem}. First, we describe the sets, indices and variables present in the formulation of the program. We then convert the constraints given in Table \ref{tbl:constraint-summary} into mathematical terms, and outline the objective function of the linear program.

@@

\subsection{Variables} \label{sec:meth-variables}
Since each clinician may be assigned to work on any service, during any block of the year, we denote such an assignment as a binary variable $X_{c, b, s}$. A value of 1 indicates that the given clinician $c$ is assigned to provide service $s$ during block $b$. Weekend assignments are similarly defined using a binary variable $Y_{c, w}$, but without a service index, as clinicians are expected to provide all services during the weekends. We then introduce a set of constants $m_{c, s}$ and $M_{c, s}$ to constrain the minimal and maximal number of blocks each clinician is allowed to work during the year. Table \ref{tbl:variables-constants} presents a summary of the constants and variables in the problem.
%To optimize the soft constraint Block-Weekend Adjacency, we maximize the product $X_{c, b, s} \cdot Y_{c, w}$ for adjacent blocks and weekends. To formulate such an objective as a linear function of variables, we introduce another set of variables, denoted by $Z_{c, b, s}$, with additional constraints on its range. Further details regarding this variable are described in Section \ref{sec:meth-objectives}.

\begin{table}[h]
	\centering
	\begin{tabular}{ c l }
		\hline
		\textbf{Name}              & \textbf{Description}                                               \\ \hline
		$X_{c, b, s} \in \{0, 1\}$ & assignment of clinician $c$ to service $s$ on block $b$            \\
		$Y_{c, w} \in \{0, 1\}$    & assignment of clinician $c$ on weekend $w$                         \\
%		$Z_{c, b, s} \in \{0, 1\}$ & helper variable for optimizing Block-Weekend adjacency             \\
		$m_{c, s}$                 & minimum number of blocks clinician $c$ should cover on service $s$ \\
		$M_{c, s}$                 & maximum number of blocks clinician $c$ should cover on service $s$
	\end{tabular}
	\caption{Description of variables and constants in the problem}
	\label{tbl:variables-constants}
\end{table}

\subsection{Constraints} \label{sec:meth-constraints}
We now formalize the hard constraints in Table \ref{tbl:constraint-summary} using the variables defined above.

\begin{align}
	&\sum_{c=1}^{C} X_{c, b, s} = 1 &&\forall b\in \mathcal{B}, s \in \mathcal{S} \tag{BC} \label{eqn:constr-block-cov} \\
	&\sum_{c=1}^{C} Y_{c, w} = 1 &&\forall w\in \mathcal{W} \tag{WC} \label{eqn:constr-weekend-cov} \\
	&m_{c, s} \leq \sum_{b=1}^{B} X_{c, b, s} \leq M_{c, s} &&\forall\ c\in\mathcal{C}, s\in\mathcal{S} \tag{MM} \label{eqn:constr-min-max} \\
	&X_{c, b, s} + X_{c, b + 1, s} \leq 1 &&\forall c\in\mathcal{C}, b \leq B - 1, s\in\mathcal{S} \tag{NCB} \label{eqn:constr-no-consec-blocks} \\
	&Y_{c, w} + Y_{c, w + 1} \leq 1 &&\forall c\in\mathcal{C}, w \leq W - 1 \tag{NCW} \label{eqn:constr-no-consec-weekends} \\
	&\floor*{\frac{W}{C}} \leq \sum_{w=1}^W Y_{c, w} \leq \ceil*{\frac{W}{C}} &&\forall c\in\mathcal{C} \tag{EW} \label{eqn:constr-equal-weekends} \\
	&\floor*{\frac{\abs{\mathcal{L}}}{C}} \leq \sum_{w\in\mathcal{L}} Y_{c, w} \leq \ceil*{\frac{\abs{\mathcal{L}}}{C}} &&\forall c\in\mathcal{C} \tag{EH} \label{eqn:constr-equal-holidays}
\end{align}

\subsection{Objectives} \label{sec:meth-objectives}
As described in Section \ref{sec:problem}, the soft constraints of the clinician scheduling problem include: satisfying clinician block off requests (BR), satisfying clinician weekend off requests (WR), and assigning weekends closer to blocks (BWA). We convert these soft constraints into linear objective functions of the binary variables defined in Section \ref{sec:meth-variables}. Objectives (\ref{eqn:obj-block-requests}) and (\ref{eqn:obj-weekend-requests}) can be written as the following linear functions:  %minor suggestion - consider numbering the objectives for ease later when saying things like "these two objectives"...or writing the full name of the objectives instead of 'this' and 'these' as sometimes can get confusing what the 'this' and the 'these' refer to. 
\begin{align}
	&Q_1(X) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} (-1)^{\ind(b\,\in\,\mathcal{BR}_c)}\cdot X_{c, b, s} \tag{BR} \label{eqn:obj-block-requests}\\
	&Q_2(Y) = \sum_{c=1}^{C} \sum_{w=1}^{W} (-1)^{\ind(w\,\in\,\mathcal{WR}_c)}\cdot Y_{c, w} \tag{WR} \label{eqn:obj-weekend-requests}
\end{align}
where $\ind(P)$ is the indicator function that has value 1 when predicate $P$ holds and 0 otherwise. In the above two objectives, we penalize any assignments that conflict with a block or weekend request, and aim to maximize the non-conflicting assignments. \\

The Block-Weekend Adjacency is optimized by considering the product $X_{c, b, s}\cdot Y_{c, w}$ for values of $w$ ``adjacent'' to the value of $b$. This leads to the maximization objective %avoid editorial adjectives
\begin{align}
&Q_3(X, Y) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} X_{c, b, s}\cdot Y_{c, w=\varphi(b)} \tag{BWA} \label{eqn:obj-block-weekend-adj}
\end{align}
where $\varphi(b)$ maps a block one-to-one to an adjacent weekend, by some appropriate definition of adjacency. For instance, clinicians might want to be assigned during a weekend that falls within an assigned block. In this case, we will have $\varphi(b) = 2b-1$. \\

%So if clinician $c$ is assigned to work during block 3, corresponding to weeks 5 and 6 assuming 2-week blocks, they might also want to be assigned to work during weekend 5. In that case, we would like $X_{c, b=3, s} \cdot Y_{c, w=5}$ to be 1, since that indicates both variables are assigned. If at least one of the two variables is not assigned, the product will be 0. 
% I really liked the example explanation!
%For instance, in the above example we will have $\varphi(b) = 2b - 1$. \\

However, as it is, $Q_3$ is not a linear function of its variables and cannot be optimized in a linear programming framework. An approach used to convert such functions into linear objectives involves introducing a helper variable and additional constraints \cite{hammer_boolean_1968}. In our case, introducing a variable $Z_{c, b, s}$ for every product $X_{c, b, s} \cdot Y_{c, w}$ with $w = \varphi(b)$, and constraining $Z$ such that 
\begin{align}
	&Z_{c, b, s} \leq X_{c, b, s} \label{eqn:helper-x-constraint}\\
	&Z_{c, b, s} \leq Y_{c, w=\varphi(b)} &&\forall s\in\mathcal{S} \label{eqn:helper-y-constraint}
\end{align}
allows us to rewrite $Q_3$ as a linear function of $Z$,
\begin{align}
	&Q_3(Z) = \sum_{c=1}^{C} \sum_{b=1}^{B} \sum_{s=1}^{S} Z_{c, b, s}
\end{align}
Indeed from Eqns. (\ref{eqn:helper-x-constraint}) and (\ref{eqn:helper-y-constraint}),  whenever $X_{c, b, s} \cdot Y_{c, w} = 1$ (respectively, 0), $Z_{c, b, s}$ can attain a maximum value of 1 (respectively, 0), giving us the correct adjacency maximization objective. \\
%Indeed, whenever $X_{c, b, s} \cdot Y_{c, w} = 1$, $Z_{c, b, s}$ can attain a maximum value of 1, and whenever $X_{c, b, s} \cdot Y_{c, w} = 0$, at least one of $X_{c, b, s}$ or $Y_{c, w}$ must be 0, so $Z_{c, b, s}$ will be constrained to attain a maximum value of 0, giving us the correct adjacency maximization objective. \\

In order to optimize all objectives simultaneously, we optimize a weighted sum of the normalized objective functions,
\begin{equation}
\max_{X, Y, Z} \alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
\end{equation}
where $\bar{Q}_i$ is the normalization of objective $Q_i$, and $0 \leq \alpha, \beta \leq 1$. This method guarantees an optimal solution to be Pareto optimal \cite{stanimirovic_linear_2011}. \\

%Thus, our clinician scheduling problem is a multiple objective optimization problem. The most common approach to solving multiple objective optimization problems is by optimizing a weighted sum of the normalized objective functions, as this guarantees the optimal solution to be Pareto optimal [ref \ref{???}]. This is the approach we decided to use in our problem, to ensure all three objectives are considered when finding a solution. Under the assumption that each clinician in $\mathcal{C}$ provides all types of services in $\mathcal{S}$, the normalized objectives can be written as follows,
%\begin{align}
%	&\bar{Q}_1(X) = \frac{Q_1(X)}{C \cdot B \cdot S} \tag{Block Requests} \label{eqn:norm-obj-block-requests}\\
%	&\bar{Q}_2(Y) = \frac{Q_2(Y)}{C \cdot W} \tag{Weekend Requests} \label{eqn:norm-obj-weekend-requests} \\
%	&\bar{Q}_3(Z) = \frac{Q_3(Z)}{C \cdot B \cdot S} \tag{Block-Weekend Adjacency} \label{eqn:norm-obj-block-weekend-adj}
%\end{align}
%where we divide each of the original objective functions by the sum of the absolute values of its coefficients [ref \ref{???}]. The final weighted objective is given by
%\begin{equation}
%	\alpha \bar{Q}_1(X) + \beta \bar{Q}_2(Y) + (1 - \alpha - \beta) \bar{Q}_3(Z)
%\end{equation}
%with $0 \leq \alpha, \beta \leq 1$. \\

### 2019-08-29 paper/paper.problem.tex
Clean up abstract, intro, problem, methods, results
Make methods section more concise + add missing references

Several constraints are placed on the on-call assignments. First, each clinician has limits on the number of blocks they can and must work during the year, depending on the type of consultation. For instance, a clinician might have to provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV consultation throughout the year. These limits may change from year to year as the number of clinicians in the department changes. Moreover, the schedule should not assign a clinician to work for two blocks or two weekends in a row. The schedule should also distribute both regular and holiday weekends equally among all clinicians. \\

@@

\begin{table}[h]
	\centering
	\begin{tabular}{ l l l }
		\hline
		\textbf{Constraint Name} & \textbf{Description}                                                                                                           & \textbf{Type} \\ \hline
		Block Coverage           & \makecell[l]{each service needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
		Weekend Coverage         & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
		Min/Max                  & \makecell[l]{for a given service, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
		No Consecutive Blocks    & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all services}                                    & Hard          \\ \hline
		No Consecutive Weekends  & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline
		Equal Weekends           & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     & Hard          \\ \hline
		Equal Holidays           & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                & Hard          \\ \hline
		Block Requests           & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        & Soft          \\ \hline
		Weekend Requests         & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      & Soft          \\ \hline
		Block-Weekend Adjacency  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}                                     & Soft
	\end{tabular}
	\caption{Summary of the constraints for the clinician scheduling problem}
	\label{tbl:constraint-summary}
\end{table}

In most scheduling problems, the constraints can be divided into hard and soft constraints. Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions [ref]. Typically, soft constraints are encoded as objective functions rather than actual constraints. These objective functions are maximized or minimized when finding a solution. In the case of our clinician scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend Adjacency as our soft constraints, while the rest of the constraints are hard. Though it is important to take clinician requests into account when constructing the schedule, it is crucial that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled. \\

It is important to note that - irrespective of how the schedule is generated - clinicians may exchange certain weeks or days throughout the year after the schedule is implemented. The approach to solving the clinician scheduling problem does not account for these future exchanges, and only focuses on the full year time horizon.

>>>>

Several constraints are placed on the on-call assignments. First, each clinician has limits on the number of blocks they can and must work during the year, depending on the type of consultation. For instance, a clinician might have to provide 3-5 blocks of general ID consultation as well as 2-3 blocks of HIV consultation throughout the year. These limits may change from year to year as the number of clinicians in the department changes. Moreover, the schedule should not assign a clinician to work for two consecutive blocks or two consecutive weekends in a row. The schedule should also distribute both regular and holiday weekends equally among all clinicians. \\

@@

\begin{table}[h]
	\centering
	\begin{tabular}{ l l l }
		\hline
		\textbf{Constraint Name} & \textbf{Description}                                                                                                           & \textbf{Type} \\ \hline
		Block Coverage (BC)           & \makecell[l]{each service needs to have exactly \\ one clinician that covers any given block}                                 & Hard          \\ \hline
		Weekend Coverage (WC)        & \makecell[l]{every weekend needs to have exactly \\ one clinician that covers it}                                              & Hard          \\ \hline
		Min/Max (MM)                  & \makecell[l]{for a given service, each clinician can only \\ work between the minimum and maximum \\ number of allowed blocks} & Hard          \\ \hline
		No Consecutive Blocks (NCB)   & \makecell[l]{any clinician should not work \\ two consecutive blocks, across all services}                                    & Hard          \\ \hline
		No Consecutive Weekends (NCW) & \makecell[l]{any clinician should not work two consecutive \\ weekends}                                                        & Hard          \\ \hline
		Equal Weekends (EW)          & \makecell[l]{weekends should be equally distributed \\ between clinicians}                                                     & Hard          \\ \hline
		Equal Holidays (EH)          & \makecell[l]{long weekends should be equally distributed \\ between clinicians}                                                & Hard          \\ \hline
		Block Requests (BR)         & \makecell[l]{each clinician can request to be off service \\ during certain blocks throughout the year}                        & Soft          \\ \hline
		Weekend Requests (WR)        & \makecell[l]{each clinician can request to be off service \\ during certain weekends throughout the year}                      & Soft          \\ \hline
		Block-Weekend Adjacency (BWA)  & \makecell[l]{the block and weekend assignments of a given \\ clinician should be adjacent}                                     & Soft
	\end{tabular}
	\caption{Summary of the constraints for the clinician scheduling problem}
	\label{tbl:constraint-summary}
\end{table}

In most scheduling problems, the constraints can be divided into hard and soft constraints. Hard constraints must be satisfied by any candidate solution, while soft constraints can be used to select a more favourable solution from all candidate solutions. Typically, soft constraints are encoded as objective functions rather than actual constraints [ref]. These objective functions are maximized or minimized when finding a solution. In the case of our clinician scheduling problem, we chose Block Requests, Weekend Requests and Block-Weekend Adjacency as our soft constraints, while the rest of the constraints are hard. Though it is important to take clinician requests into account when constructing the schedule, it is crucial that the work-load of the schedule is balanced among all clinicians, and the needs of the patients are fulfilled. It is important to note that - irrespective of how the schedule is generated - clinicians may exchange certain weeks or days throughout the year after the schedule is implemented. The approach to solving the clinician scheduling problem does not account for these future exchanges, and only focuses on the full time horizon.
